---
title: "Simple document"
author: Yuanyuan Zeng
data: 2021-09-16
output: github_document
---
```{r, echo = FALSE, message = FALSE, warming = FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```

## Problem 1

```{r, message = FALSE, warming = FALSE}
# Read and clean the Mr.Trash Wheel sheet
trash_data = 
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N535") %>%
  janitor::clean_names() %>% 
  filter(dumpster != "NA") %>% 
  mutate(
    sports_balls = round(sports_balls))

head(trash_data, 5)
```

```{r, message = FALSE, warming = FALSE}
# Read and clean the precipitation data for 2018
pre_2018 = (
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
    janitor::clean_names() %>% 
    mutate(year = 2018) %>%  
    relocate(year))
head(pre_2018, 5)

# Calculate total precipitation in 2018
sum(pull(pre_2018,var =3))
```

```{r, message = FALSE, warming = FALSE}
# Read and clean the precipitation data for 2019
pre_2019 = (
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2019 Precipitation", range = "A2:B14") %>% 
    janitor::clean_names() %>% 
    mutate(year = 2019) %>% 
    relocate(year)
  )

head(pre_2019, 5)

# Combine two precipitation data
precipitation_data = (
  bind_rows(pre_2018, pre_2019) %>% 
  mutate(month = month.name[as.numeric(month)])
)

head(precipitation_data, 5)

# find the median of number of sports ball in dumpster in 2019
filter(trash_data, year == "2019") %>% 
  pull(sports_balls) %>% 
  median()
```

* The trash_data contains 488 rows and 14 columns. There are 14 variables which are dumpster, month, year, data, weight_tons, volume_cubic_yards, plastic_bottles, polystyrene, cigarette_butts, grocery_bags, chip_bags, sports_balls, and homes_powered. Dumpster and month are character variables. Others are numeric variables. The tables both pre_2018 and pre_2019 contains 12 rows and 3 colors. There are three numeric variables which are year, month, and total. Combining two precipitation data, we get a table named precipitation_data which contains 24 rows and 3 columns. The variable name are the same as two separated tables. The total precipitation in 2018 is 70.33 inches. The median number of sports balls in dumspter in 2019 is 9.

## Problem 2

```{r, message = FALSE, warming = FALSE}
# Clean the data in pols-month.csv
pols_month = 
  read_csv("./data/pols-month.csv") %>% 
  separate(mon,  into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    month = month.name[as.numeric(month)]) %>% 
  pivot_longer(
    prez_gop | prez_dem,
    names_to = "president",
    names_prefix = "prez_",
    values_to = "gop_or_dem") %>% 
  filter( gop_or_dem!=0 ) %>% 
  select(!day & !gop_or_dem) %>% 
  relocate(year, month, president, everything())

head(pols_month, 5)

# Clean the data in snp.csv
snp_df = 
  read_csv("./data/snp.csv") %>% 
  mutate(
    date = mdy(date)
  ) %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    year = ifelse(as.numeric(year)>2049, as.numeric(year)-100, year)
  ) %>% 
  mutate(
    month = month.name[as.numeric(month)]
    ) %>% 
  select(!day) %>% 
  relocate(year)

head(snp_df, 5)

# Clean th unemployment data 
unemployment_df = 
  read_csv("./data/unemployment.csv", col_types = "cdddddddddddd") %>% 
  pivot_longer(
    Jan : Dec,
    names_to = "month",
    values_to = "unemployment"
  ) %>% 
  filter(unemployment !="NA") %>% 
  janitor::clean_names() %>% 
  mutate(
    month = match(month, month.abb),
    month = month.name[as.numeric(month)]
  ) %>% 
  arrange(desc(year)) 

head(unemployment_df, 5)

# left join snp into pols data set
pols_snp =
  left_join(pols_month, snp_df, by = c("year","month"))

# left join unemployment into pols_snp 
pols_snp_unemp =
  left_join(pols_snp, unemployment_df, by = c("year", "month"))

# find the range of year in the data set
range(pull(pols_snp,var =1))
```

* In the pols_month data set, there are 822 rows and 9 variables. The year, month, and president are character variables. The column of president show either dem or gop. The data set of snp_df contains 787 rows and 3 variables which are year, month and close. The close represents the closing values of the S&P stock index. The data set of unemployment_df contains 810 rows and 3 variables which are year, month, and unemployment. We merge the snp_df into lops_month by using left join which means that we keep everything in lops_month dataset and add closing values of stock into it. The resulting data set is pols_snp, which contains 822 rows and 10 variables. The year ranges from 1947 to 2015. The resulting data set of merging unemployment into pols_snp contains same number of rows as before and have one additional variable called unemployment.There are 822 rows and 11 variables. There are missing values in column of close and unemployment, which means that closing value of stock and unemployment are not available to the corresponding year and month.

## Problem 3

```{r, message = FALSE, warming = FALSE}
# Clean the data on popularity of baby names
baby_names = 
  read.csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>%
  distinct() %>% 
  mutate(
    gender = tolower(gender),
    ethnicity = tolower(ethnicity),
    ethnicity = replace(ethnicity, ethnicity == "asian and paci", "asian and pacific islander"),
    ethnicity = replace(ethnicity, ethnicity == "black non hisp", "black non hispanic"),
    ethnicity = replace(ethnicity, ethnicity == "white non hisp", "white non hispanic"),
    child_s_first_name = tolower(child_s_first_name)
  )

head(baby_names, 5)

# Create table showing popularity of the name "Olivia"
olivia_df =
  baby_names %>% 
  select(-count) %>% 
  filter(child_s_first_name == "olivia") %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  )

# Create table showing popularity of the most popular name among male children over time
popular_names =
  baby_names %>% 
  select(-count) %>% 
  filter(
    gender != "female",
    rank == "1"
  ) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "child_s_first_name"
  )

# Produce scatter plot
male_non_his =
  baby_names %>% 
  filter(
    gender != "female",
    year_of_birth == "2016",
    ethnicity == "white non hispanic"
  )

ggplot(male_non_his, aes(x= rank, y = count))+
  geom_point() +
  labs(
    title = "popularity of white non hispanic male children's names",
    x = "rank in popularity of the name",
    y = "number of children with the name"
  )
```


